Analysing Constraint Grammars with a SAT-solver

### 2016-03-08 paper/lrec/eval_futurework.tex
cleanup & more examples to future work

% These two rules both remove an adverb, but the first one has a broader condition:
% remove adverb if followed by a noun. In contrast, the second rule has a stricter condition: only remove adverb if it is preceded by a determiner, the adverb itself is ambiguous with an adjective, and followed by a noun. 
The problem is that that
The first rule has a broader condition, 
it removes the adverb in all possible cases 
 removes the adverb in all possible cases, and the second will not have any chance to act. If the rules were in the opposite order, then there would be no problem.

%%% Nevermind, this wasn't a conflict, my program was just buggy :-P

%\begin{itemize}
%\item[] 
%\begin{verbatim}
%SELECT DetPosNotZijn IF (1 Noun) ;
%SELECT DetPos IF (-1 (vbser pres p3 sg)) 
%                 (0 "zijn") (1 Noun);
%\end{verbatim}
%\end{itemize}

% SELECT DetPosNotZijn IF (1 Noun);
%If the first rule fires, then the condition of the second rule cannot be true: there can never be a ``zijn'' in the target, because the previous rule has selected everything that is a possessive determiner and not ``zijn''. 
%So, the first rule must not fire. It cannot be because condition does not hold: the condition (1 Noun) is shared with the second rule, and

@@

\subsection{Spanish} The Spanish grammar had proportionately the
highest number of errors. The grammar we ran is like the one
found in the Apertium repository (linked on the previous page), 
apart from two changes: we fixed some typos (capital O for 0) in order to make it compile, and
commented out two rules that used regular expressions, because our
solution does not handle them properly; see more on Section~\ref{sec:conclusion}.

We include two versions of the Spanish grammar in table~\ref{table:res}: in column \textsc{spa}, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column \textsc{spa}~{\small sep.~lem.}, we just added each word form and lemma as individual readings, allowed to combine with any other reading. 
This latter version ran much faster, but failed to detect two internal conflicts, both for rules that target a lemma in a condition.
% SELECT:pr_cnjadv_3 cnjadv IF (0 "después de") 
More interesting observation, the first version failed to detect a true interaction conflict, because it marked the two rules as internal conflicts.
%In addition, there were differences in the generated output; examples will be shown later in this section.

@@

\begin{table*}[t!]
\centering
\begin{tabular}{|p{2.5cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}|}

\hline
              & 1 clitic + 
                lemmas from 
                 grammar & 2 clitics 
                           + lemmas from 
                              grammar & 1 clitic +
                                          all readings 
                                          from grammar    & all readings from grammar \\ \hline
\# readings   & 5851 (4263 
                 + 1588) & 9494
                       (7906 + 1588) & 6657 (4263 + 2394) & 2394  \\ \hline
\# conflicts  & 214      & 214       & 22                & 22  \\ \hline
\# internal/interaction & 211 + 3    & 211 + 3           & 19 + 3    & 19 + 3  \\ \hline

\# true 
   positives  & n        & m         & o                & p  \\ \hline
\# false 
   positives  & p        & q         & r                & s   \\ \hline
\clock{}all 
       rules & 12min     & 23min     & 12min &  4min 40s \\ \hline

@@

The rules above have been simplified to show only the relevant part.
The conflict lies in the fact that \texttt{Comma} is a subset of \texttt{Comma/N/Pron/Q}:
there is no way to trigger the second without placing a comma in position -1, and thereby triggering the first rule.

% All in all, using readings from grammar directly may be useful in order to find just the , sometimes you can have a typo, like in \emph{l\"{a}hetii}:

% \begin{verbatim}
% "w2"
%         w2<"se"><ela>
%         w2<"<lähetii>"><vblex>
% \end{verbatim}

% The second part concerns the tools in use, specifically the Apertium morphological lexicon. We noticed, only after running the experiments, that the expansion of the lexicon does not work properly: for instance, some nominal forms of the verbs were reported as nouns instead of verbs.
% We reported the errors to the developers of the lexicon, but as there was no time to fix them, we used the readings that we got from expanding the lexicon, and manually added either noun, verb or adjective tag where it was missing.

% kosimisiin:kosia<n><pl><ill>
% kosimattomuudekseen:kosia<sg><tra><pxsp3>

% Secondly, we added ambiguity class constraints for only
% \emph{intraparadigmatic ambiguities}: homonymous wordforms belonging
% to the same lexeme, in a way that is systematic to the whole
% paradigm. We reduced the morphological lexicon to contain only one
% example of each inflection paradigm, and then expanded it: with all
% the clitics, this resulted in \todo{n} word forms. Then we applied the
% normal method of creating the ambiguity classes, and specified some
% default cases manually, such as punctuation and word boundaries.

@@

This may also be feasible to support regular expressions, in a nicer
way than just spitting out the regex. (INARI: You said MORE ON THAT here, good because I don't understand what you want to say here :)

@@

\paragraph{Support for grammar writers}
One additional feature is to suggest reformattings for a rule. Recall
figure~\ref{fig:regroup} from the introduction; in that case, the
original rule was written by the original author, and another
grammarian thought that the latter form is nicer to read. Doing the
reverse operation could also be possible. If a rule with long
disjunctions conflicts, it may be useful to split it into smaller
conditions, and eliminate one at a time, in order to find the
reason(s) for the conflict.
Our next step is to evaluate our tools together with actual grammar writers,
in comparison with a corpus-based method or machine learning.

>>>>

% These two rules both remove an adverb, but the first one has a broader condition:
% remove adverb if followed by a noun. In contrast, the second rule has a stricter condition: only remove adverb if it is preceded by a determiner, the adverb itself is ambiguous with an adjective, and followed by a noun. 
The problem is that the first rule has a broader condition than the second.
The first rule removes the adverb in all possible cases where the second would, and the second will not have any chance to act. 
If the rules were in the opposite order, then there would be no problem.

@@

\subsection{Spanish} The Spanish grammar had proportionately the
highest number of errors. The grammar we ran is like the one
found in the Apertium repository (linked on the previous page), 
apart from two changes: we fixed some typos (capital O for 0) in order to make it compile, and
commented out two rules that used regular expressions, because we did not implement the support for them yet.

We include two versions of the Spanish grammar in table~\ref{table:res}: in column \textsc{spa}, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column \textsc{spa}~{\small sep.~lem.}, we just added each word form and lemma as individual readings, allowed to combine with any other reading. 
This latter version ran much faster, but failed to detect an internal conflict for a rule that targets a lemma in a condition.
% SELECT:pr_cnjadv_3 cnjadv IF (0 "después de") 
Surprisingly, the first version failed to detect a true interaction conflict, because it marked the two rules both as internal conflicts.

@@

\begin{table*}[t!]
\centering
\begin{tabular}{|p{3.15cm}|p{2.55cm}|p{2.55cm}|p{2.7cm}|p{2.1cm}|}

\hline
              & 1 clitic + 
                lemmas from 
                 grammar & 2 clitics 
                           + lemmas from 
                              grammar & 1 clitic +
                                          all readings 
                                          from grammar    & all readings from grammar \\ \hline
\# readings   & 5851 (4263~+~1588)
                       & 9494
                       (7906~+~1588) & 6657 (4263~+~2394) & 2394  \\ \hline
\#~conflicts  & 214    & 214         & 22                 & 22 \\
(internal + 
 interaction) & (211~+~3) & (211~+~3) & (19~+~3)        &  (19~+~3)  \\ \hline

\clock{} all 
       rules & Xh 12min     & Xh 23min   & Xh 12min &  Xh 4min \\ \hline

@@

The rules above have been simplified to show only the relevant part.
The conflict lies in the fact that \texttt{Comma} is a subset of \texttt{Comma/N/Pron/Q}:
there is no way to trigger the second rule without placing a comma in position -1, and thereby triggering the first rule.

@@

@@

\paragraph{Support for grammar writers}
As mentioned earlier, we could support additional checks for common issues,
such as conditions that concern the target word.
Another possible feature is to suggest reformattings for a rule. Recall
figure~\ref{fig:regroup} from the introduction; in that case, the
original rule was written by the original author, and another
grammarian thought that the latter form is nicer to read. Doing the
reverse operation could also be possible. If a rule with long
disjunctions conflicts, it may be useful to split it into smaller
conditions, and eliminate one at a time, in order to find the
reason(s) for the conflict.
Our next step is to evaluate our tools together with actual grammar writers,
in comparison with a corpus-based method or machine learning.
